---
jupyter:
  jupytext:
    formats: ipynb,md,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.1'
      jupytext_version: 1.2.1
  kernelspec:
    display_name: R 3.6.1
    language: R
    name: ir361
---

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
# Incident variance 2020-08-03
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
## Setup
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
library(oildata)
library(tidyverse)
library(stringr)
library(tidytext)
library(topicmodels)
library(ldatuning)
library(SnowballC)
library(DT)
```

```{r slideshow={'slide_type': 'subslide'}}
head(incidents)
```

```{r}
names <- unique(word(incidents$name, 1))
names
```

```{r}
df <- subset(incidents, on_offshore == "onshore" & commodity %in% c("rpp", "crude") & year >= 2002 & year < 2020)
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Narratives: properties
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### NAs
<!-- #endregion -->

```{r}
nrow(df)
nrow(subset(df, is.na(narrative)))
nrow(subset(df, significant == T))
nrow(subset(df, significant == T & is.na(narrative)))
```

For significant incidents, lack of narrative is not too concerning

```{r slideshow={'slide_type': 'subslide'}}
df <- subset(df, !is.na(narrative))
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Cleanliness
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
head(select(df, incident_ID, narrative))
```

```{r slideshow={'slide_type': 'subslide'}}
print(subset(df, incident_ID == 20020032)$narrative)
```

Quality seems ok.

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### How long are the narratives?
<!-- #endregion -->

```{r}
table(round(str_length(df$narrative), -2))
```

```{r slideshow={'slide_type': 'subslide'}}
df %>%
    mutate(len = str_length(narrative)) %>%
    ggplot(aes(x = len)) +
        geom_histogram()
```

```{r slideshow={'slide_type': 'subslide'}}
df %>%
    filter(significant == T) %>%
    mutate(len = str_length(narrative)) %>%
    ggplot(aes(x = len)) +
        geom_histogram()
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Do the narratives get longer over time?
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
df %>%
    mutate(len = str_count(narrative, "\\S+")) %>%
    filter(year >= 2002, year < 2020) %>%
    ggplot(aes(x = factor(year), y = len)) + 
        geom_boxplot(notch = T, outlier.shape = NA) + 
        scale_y_continuous(limits=c(0, 350), oob=scales::rescale_none, breaks = seq(0, 350, 50))
```

```{r slideshow={'slide_type': 'subslide'}}
df %>%
    mutate(len = str_count(narrative, "\\S+")) %>%
    filter(year >= 2002, year < 2020, significant == T) %>%
    ggplot(aes(x = factor(year), y = len)) + 
        geom_boxplot(notch = T, outlier.shape = NA) + 
        scale_y_continuous(limits=c(0, 450), oob=scales::rescale_none, breaks = seq(0, 550, 50))
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Simple topic model
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Prepare
<!-- #endregion -->

```{r}
narratives <- select(df, incident_ID, narrative)

head(narratives, 3)
```

```{r}
word_counts <- narratives %>%
    unnest_tokens(word, narrative) %>%
    anti_join(data.frame(word = str_to_lower(names)), by = "word") %>%
    anti_join(stop_words, by = "word") %>%
    filter(! str_detect(word, "^[0-9]")) %>%
    mutate(word = wordStem(word)) %>%
    count(incident_ID, word, sort = T) %>%
    cast_dtm(document = incident_ID, term = word, 
             value = n)
word_counts
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
### Find number of topics
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
# results_1 <- FindTopicsNumber(word_counts, 
#                               topics = c(5, 10, 20, 40), 
#                               metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
#                               method = "Gibbs", 
#                               control = list(seed = 532))
# saveRDS(results_1, file = ".temp/results_1.rds")
results_1 <- readRDS(".temp/results_1.rds")
```

```{r slideshow={'slide_type': 'subslide'}}
FindTopicsNumber_plot(results_1)
```

```{r slideshow={'slide_type': 'subslide'}}
# results_2 <- FindTopicsNumber(word_counts, 
#                             topics = c(15, 18, 19, 20, 21, 22, 25), 
#                             metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
#                             method = "Gibbs", 
#                             control = list(seed = 544))
# saveRDS(results_2, file = ".temp/results_2.rds")
results_2 <- readRDS(".temp/results_2.rds")
```

```{r}
FindTopicsNumber_plot(results_2)
```

```{r slideshow={'slide_type': 'subslide'}}
# results_3 <- FindTopicsNumber(word_counts, 
#                             topics = seq(15, 35), 
#                             metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
#                             method = "Gibbs", 
#                             control = list(seed = 912))
# saveRDS(results_3, file = ".temp/results_3.rds")
results_3 <- readRDS(".temp/results_3.rds")
```

```{r}
FindTopicsNumber_plot(results_3)
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
### Inspect model
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
# narratives_lda <- LDA(word_counts, method = "Gibbs", k = 25, control = list(seed = 912))
# saveRDS(narratives_lda, ".temp/narratives_lda.rds")
narratives_lda <- readRDS(".temp/narratives_lda.rds")
narratives_lda
```

```{r slideshow={'slide_type': 'subslide'}}
betas <- tidy(narratives_lda, matrix = "beta")
head(betas)
```

```{r slideshow={'slide_type': 'subslide'}}
top_terms <- betas %>%
    arrange(desc(beta)) %>%
    group_by(topic) %>%
    slice_head(n = 5) %>%
    ungroup() %>%
    arrange(topic, -beta)

head(top_terms, 10)
```

```{r slideshow={'slide_type': 'subslide'}}
# datatable(top_terms, options = list(pageLength = 5))
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Look at terms and example for topics
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Setup
<!-- #endregion -->

```{r}
gammas <- tidy(narratives_lda, matrix = "gamma")
head(gammas)
```

```{r slideshow={'slide_type': 'subslide'}}
narratives <- select(df, incident_ID, narrative)
narratives$incident_ID <- as.character(narratives$incident_ID)

get_narrative <- function(topic) {
    gammas %>%
        filter(topic == {{topic}}) %>%
        filter(gamma == max(gamma)) %>%
        left_join(narratives, by = c("document" = "incident_ID")) %>%
        {.$narrative}
}

get_terms <- function(topic) {
    top_terms %>%
        filter(topic == {{topic}}) %>%
        mutate(term = reorder_within(term, beta, topic)) %>%
        ggplot(aes(term, beta)) +
            geom_bar(stat = "identity") +
            scale_x_reordered()
}
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
This is where we will store incident cause type topics.
<!-- #endregion -->

```{r}
causes <- c()
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Topic 1
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
get_terms(topic = 1)
```

```{r}
get_narrative(topic = 1)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
This is a good candidate for a narrative that describes incident causes.
<!-- #endregion -->

```{r}
causes <- c(causes, 1)
print(causes)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Topic 2
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
get_terms(2)
```

```{r slideshow={'slide_type': 'subslide'}}
get_narrative(2)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
This topic is more likely describing the emergency response.
<!-- #endregion -->

### Topic 3

```{r slideshow={'slide_type': 'subslide'}}
get_terms(3)
```

```{r slideshow={'slide_type': 'subslide'}}
get_narrative(3)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
This topic seems to capture contractor mistakes.
<!-- #endregion -->

```{r}
causes <- c(causes, 3)
causes
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Topic 4
<!-- #endregion -->

```{r}
get_terms(4)
```

```{r slideshow={'slide_type': 'subslide'}}
get_narrative(4)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
This topic also seems to describe a response type.
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Topic 5
<!-- #endregion -->

```{r}
get_terms(5)
```

```{r slideshow={'slide_type': 'subslide'}}
get_narrative(5)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
Again, this seems to describe a failure type.
<!-- #endregion -->

```{r}
causes <- c(causes, 5)
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## See development of causes over time
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
df$incident_ID <- as.character(df$incident_ID)

topic_by_year <- gammas %>%
    left_join(select(df, incident_ID, year), by = c("document" = "incident_ID")) %>%
    group_by(year, topic) %>%
    summarize(gamma = sum(gamma) / n()) %>%
    ungroup()
head(topic_by_year)
```

```{r slideshow={'slide_type': 'subslide'}}
topic_development <- topic_by_year %>%
    filter(topic %in% causes) %>%
    ggplot(aes(x=year, y=gamma)) +
        geom_line() +
        geom_point() +
#         geom_smooth(method = "lm") +
        facet_wrap(~ topic) + 
        scale_x_continuous(minor_breaks = seq(2002, 2019, 1)) + 
        theme(panel.grid.major.x = element_line(color = "grey35", size = 0.6), 
              panel.grid.minor.x = element_line(color = "grey35", size = 0.2))
```

```{r slideshow={'slide_type': 'subslide'}}
topic_development
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Other
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
top_terms %>%
    filter(topic == 1) %>%
    mutate(term = reorder_within(term, beta, topic)) %>%
    ggplot(aes(term, beta)) +
        geom_bar(stat = "identity") +
        scale_x_reordered()
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
#### Check some narratives
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
set.seed(500)
some_narratives <- sample(gammas$document, 9)
some_narratives
```

```{r slideshow={'slide_type': 'subslide'}}
gammas %>%
    filter(document %in% some_narratives) %>%
    ggplot(aes(x=topic, y=gamma)) +
        facet_wrap(~ document) +
        geom_bar(stat="identity")
```

```{r slideshow={'slide_type': 'subslide'}}
df$incident_ID <- as.character(df$incident_ID)

topic_by_year <- gammas %>%
    left_join(select(df, incident_ID, year), by = c("document" = "incident_ID")) %>%
    group_by(year, topic) %>%
    summarize(gamma = sum(gamma) / n()) %>%
    ungroup()
head(topic_by_year)
```

```{r slideshow={'slide_type': 'subslide'}}
topic_by_year %>%
    ggplot(aes(x=year, y=gamma)) +
        geom_line() +
        geom_point() +
        geom_smooth(method = "lm") +
        facet_wrap(~ topic)
```
