---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.1'
      jupytext_version: 1.2.1
  kernelspec:
    display_name: R 3.6.1
    language: R
    name: ir361
---

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
# Research Progress Meeting 2020-02-28

## Where do I come from, where do I go?
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
* Research series presentation went well
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* But I lost track of what was said and what I was going to do
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* And I want to get stuff done
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Setup
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
library(tidyverse)
library(oildata)
library(DataAnalysisTools)
options(repr.plot.width=7, repr.plot.height=4)
```

```{r slideshow={'slide_type': 'subslide'}}
head(pipelines)
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Throwback: Learning insight from PhD Research Series
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
learning_data <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              incidents = sum(significant_incidents, na.rm = T), 
              incident_rate = incidents/miles)

head(learning_data) %>%
    jupyter_styling(font_size = 14)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Incident counts
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = incidents)) +
    geom_col()
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Pipeline network expansion
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = miles)) +
    geom_line() + 
    scale_y_continuous(limits = c(0, 220000))
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Incident rate over time
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = incident_rate)) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Incident rate by commodity
<!-- #endregion -->

```{r}
learning_by_commodity <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year, commodity) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              incidents = sum(significant_incidents, na.rm = T), 
              incident_rate = incidents/miles)

head(learning_by_commodity) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
ggplot(learning_by_commodity, aes(x = year, y = incident_rate, color = commodity)) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Learning of largest actors
<!-- #endregion -->

```{r}
largest_companies <- pipelines %>%
    group_by(year, ID) %>%
    summarize(total_miles = sum(total_miles, na.rm = T)) %>%
    group_by(ID) %>%
    filter(total_miles == max(total_miles)) %>%
    ungroup() %>%
    top_n(6, total_miles)

largest_companies %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company <- pipelines %>%
    filter(ID %in% largest_companies$ID) %>%
    group_by(year, ID) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              incidents = sum(significant_incidents, na.rm = T), 
              incident_rate = incidents/miles) %>%
    ungroup() %>%
    mutate(name = add_company_names(ID))

head(learning_by_company) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = incidents)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = incident_rate)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Learning: Gallons spilled
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
learning_data <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              volume = sum(incidents_volume, na.rm = T), 
              volume_per_mile = volume/miles)

head(learning_data) %>%
    jupyter_styling(font_size = 14)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Total gallons spilled
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = volume)) +
    geom_col()
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Gallons spilled, normalized trend
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = volume_per_mile)) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Gallons spilled, by commodity
<!-- #endregion -->

```{r}
learning_by_commodity <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year, commodity) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              volume = sum(incidents_volume, na.rm = T), 
              volume_per_mile = volume/miles)

head(learning_by_commodity) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
ggplot(learning_by_commodity, aes(x = year, y = volume_per_mile)) +
    facet_wrap(~ commodity) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Gallons spilled, learning by largest companies
<!-- #endregion -->

```{r}
learning_by_company <- pipelines %>%
    filter(ID %in% largest_companies$ID) %>%
    group_by(year, ID) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              volume = sum(incidents_volume, na.rm = T), 
              volume_per_mile = volume/miles) %>%
    ungroup() %>%
    mutate(name = add_company_names(ID))

head(learning_by_company) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = volume)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = volume_per_mile)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Learning: Cost
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
learning_data <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              cost = sum(incidents_cost, na.rm = T), 
              cost_per_mile = cost/miles)

head(learning_data) %>%
    jupyter_styling(font_size = 14)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Cost by year
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = cost)) +
    geom_col()
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Cost per mile, trend
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = cost_per_mile)) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Cost by commodity, trend
<!-- #endregion -->

```{r}
learning_by_commodity <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year, commodity) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              cost = sum(incidents_cost, na.rm = T), 
              cost_per_mile = cost/miles)

head(learning_by_commodity) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
ggplot(learning_by_commodity, aes(x = year, y = cost_per_mile)) +
    facet_wrap(~ commodity) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Cost for largest companies
<!-- #endregion -->

```{r}
learning_by_company <- pipelines %>%
    filter(ID %in% largest_companies$ID) %>%
    group_by(year, ID) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              cost = sum(incidents_cost, na.rm = T), 
              cost_per_mile = cost/miles) %>%
    ungroup() %>%
    mutate(name = add_company_names(ID))

head(learning_by_company) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = cost)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = cost_per_mile)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```
