---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.1'
      jupytext_version: 1.2.1
  kernelspec:
    display_name: R 3.6.1
    language: R
    name: ir361
---

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
# Research Progress Meeting 2020-02-28

## Where do I come from, where do I go?
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
* Research series presentation went well
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* But I lost track of what was said and what I was going to do
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* And I want to get stuff done
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Setup
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
library(tidyverse)
library(oildata)
library(DataAnalysisTools)
options(repr.plot.width=7, repr.plot.height=4)
```

```{r slideshow={'slide_type': 'subslide'}}
head(pipelines)
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Throwback: Learning insight from PhD Research Series
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
learning_data <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              incidents = sum(significant_incidents, na.rm = T), 
              incident_rate = incidents/miles)

head(learning_data) %>%
    jupyter_styling(font_size = 14)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Incident counts
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = incidents)) +
    geom_col()
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Pipeline network expansion
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = miles)) +
    geom_line() + 
    scale_y_continuous(limits = c(0, 220000))
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Incident rate over time
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = incident_rate)) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Incident rate by commodity
<!-- #endregion -->

```{r}
learning_by_commodity <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year, commodity) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              incidents = sum(significant_incidents, na.rm = T), 
              incident_rate = incidents/miles)

head(learning_by_commodity) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
ggplot(learning_by_commodity, aes(x = year, y = incident_rate, color = commodity)) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Learning of largest actors
<!-- #endregion -->

```{r}
largest_companies <- pipelines %>%
    group_by(year, ID) %>%
    summarize(total_miles = sum(total_miles, na.rm = T)) %>%
    group_by(ID) %>%
    filter(total_miles == max(total_miles)) %>%
    ungroup() %>%
    top_n(6, total_miles)

largest_companies %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company <- pipelines %>%
    filter(ID %in% largest_companies$ID) %>%
    group_by(year, ID) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              incidents = sum(significant_incidents, na.rm = T), 
              incident_rate = incidents/miles) %>%
    ungroup() %>%
    mutate(name = add_company_names(ID))

head(learning_by_company) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = incidents)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = incident_rate)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Learning: Gallons spilled
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
learning_data <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              volume = sum(incidents_volume, na.rm = T), 
              volume_per_mile = volume/miles)

head(learning_data) %>%
    jupyter_styling(font_size = 14)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Total gallons spilled
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = volume)) +
    geom_col()
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Gallons spilled, normalized trend
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = volume_per_mile)) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Gallons spilled, by commodity
<!-- #endregion -->

```{r}
learning_by_commodity <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year, commodity) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              volume = sum(incidents_volume, na.rm = T), 
              volume_per_mile = volume/miles)

head(learning_by_commodity) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
ggplot(learning_by_commodity, aes(x = year, y = volume_per_mile)) +
    facet_wrap(~ commodity) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Gallons spilled, learning by largest companies
<!-- #endregion -->

```{r}
learning_by_company <- pipelines %>%
    filter(ID %in% largest_companies$ID) %>%
    group_by(year, ID) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              volume = sum(incidents_volume, na.rm = T), 
              volume_per_mile = volume/miles) %>%
    ungroup() %>%
    mutate(name = add_company_names(ID))

head(learning_by_company) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = volume)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = volume_per_mile)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Learning: Cost
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
learning_data <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              cost = sum(incidents_cost, na.rm = T), 
              cost_per_mile = cost/miles)

head(learning_data) %>%
    jupyter_styling(font_size = 14)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Cost by year
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = cost)) +
    geom_col()
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Cost per mile, trend
<!-- #endregion -->

```{r}
ggplot(learning_data, aes(x = year, y = cost_per_mile)) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Cost by commodity, trend
<!-- #endregion -->

```{r}
learning_by_commodity <- pipelines %>%
    filter(commodity %in% c("crude", "hvl", "non_hvl")) %>%
    group_by(year, commodity) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              cost = sum(incidents_cost, na.rm = T), 
              cost_per_mile = cost/miles)

head(learning_by_commodity) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
ggplot(learning_by_commodity, aes(x = year, y = cost_per_mile)) +
    facet_wrap(~ commodity) +
    geom_line() +
    geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Cost for largest companies
<!-- #endregion -->

```{r}
learning_by_company <- pipelines %>%
    filter(ID %in% largest_companies$ID) %>%
    group_by(year, ID) %>%
    summarize(miles = sum(total_miles, na.rm = T), 
              cost = sum(incidents_cost, na.rm = T), 
              cost_per_mile = cost/miles) %>%
    ungroup() %>%
    mutate(name = add_company_names(ID))

head(learning_by_company) %>%
    jupyter_styling(font_size = 14)
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = cost)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

```{r slideshow={'slide_type': 'subslide'}}
learning_by_company %>%
    filter(miles > 100) %>%
    ggplot(aes(x = year, y = cost_per_mile)) +
        facet_wrap(~ name) +
        geom_line() +
        geom_smooth(method="lm")
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Notes
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Off/onshore
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* Trends (I believe) driven by offshore spills
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* Offshore pipelines are still at the beginning of their learning curve
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* (And on average spill more, because source often cannot be "shut off"?)
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### "Agenda"
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* What do I want?
    * I do not want to "kiss up" to oil companies and show their progress
    * But I also do not want to be "unfair"
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* I want to illuminate how learning progresses (or not) for less essential functions of an organization
    * For offshore pipelines, learning is still important, because of criticism
    * In terms of onshore leaks, it often affects the powerless
    * Hence learning is less of a priority
    * Location choice is already briefly mentioned by Perrow
    * Often build in cheaper locations, which usually also do not increase in price as a result of pipeline construction
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
* Is pipeline security based on technological limitations or dominated by conceptualizations (e.g., "there will always be spills")?
    * Why aren't there more "zero incident" instances?
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
* Genealogy provides a potential "blend" of the two options
    * Individuals approach their work with their presumptions, but they obtain results and feed those back into society
    * Iterative process
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Feedback
<!-- #endregion -->

The genealogical approach is potentially the driving force behind what I am looking at. The two keys are truth and power. Power can make an impact on truth, but when there is a change in the rules of formation of statements of scientific truth. The propositions which are then acceptable under the new regime are then tested and can be verified or falsified by scientists. They then re-inform society.

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
In the case of oil pipelines, it is clear who the power is with. And the truth alligns with our idea of who power is with. It is remarkable, in which context spills are acceptable, and when they are not. In disadvantaged communities, they are acceptable. The oil spill in the gulf of Mexico was not permittable. This led to learning in the second case. But still, the power structure is such that oil spills in many cases are acceptable.
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
A change in power can lead to a new regime of what are acceptable propositions. Now, we see a new regime in the sciences. Nature and Science are open to sustainability researchers, and there are many who are willing to research the oil industry there; whereas before relevant research was often sponsored by oil exploration.
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
We may see an example of divergence in paths when we look at ownership. The ownership structures impact the power balance between company and external stakeholders, especially the government. Here, it is possible that the power structure makes diverging propositions permitable. For instance, sustainability might be a permitable proposition that is picked up in parts.
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "fragment"}} -->
However, the dynamic altogether might be different than in the cases that Foucault describes. Because the coupling with the (natural) environment is stronger. Or maybe that just means that hypotheses that have tested true before (the viablity of oil extraction) are not testing true anymore.
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
Also, I forgot to mention that, but I still think the EU regulation on energy-efficient light bulbs was an important moment. It shows how entranched producers were in current technology, even as the benefit of new technology were very fast. LEDs quickly went from unfeasible to universally adopted.
<!-- #endregion -->
