---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.1'
      jupytext_version: 1.2.1
  kernelspec:
    display_name: R 3.6.1
    language: R
    name: ir361
---

# Research Progress Meeting 2020-03-xx


## Let's see how spills develop in terms of other performance metrics

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Setup
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
library(tidyverse)
library(oildata)
library(DataAnalysisTools)
library(gridExtra)
```

```{r}
packageVersion("oildata")
```

```{r}
head(pipelines)
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Overview
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
pipelines_year <- pipelines %>%
    group_by(year, on_offshore) %>%
    summarize(!!! oildata::pipelines_consolidation) %>%
    pivot_longer(cols = -c(year, on_offshore), names_to = "variable", values_to = "value")

glimpse(pipelines_year)
```

```{r slideshow={'slide_type': 'subslide'}}
unique(pipelines_year$variable)
```

```{r slideshow={'slide_type': 'subslide'}}
overview <- pipelines_year %>%
    filter(variable %in% c("miles", "significant_incidents", "incidents_cost", "incidents_volume")) %>%
    ggplot(aes(x = year, y = value)) +
        facet_wrap(on_offshore ~ variable, scales = "free") +
        geom_line()

options(repr.plot.width=8, repr.plot.height=6)
```

```{r slideshow={'slide_type': 'subslide'}}
overview
```

All cost in 1984 dollars. Volume spilled in gallons.

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### What is that cost spike in 2009 for *onshore*?!
<!-- #endregion -->

```{r}
subset(incidents, on_offshore == "onshore" & cost_1984 > 1e+8, -narrative)
```

It is the Kalamazoo River oil spill.

https://en.wikipedia.org/wiki/Kalamazoo_River_oil_spill

```{r slideshow={'slide_type': 'subslide'}}
subset(incidents, round(cost_1984) == 460446182)$narrative %>%
    jupyter_styling(font_size = 14)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### What happened with offshore in '04?
<!-- #endregion -->

```{r}
subset(incidents, on_offshore == "offshore" & cost_1984 > 1e+7, -c(narrative, DATAFILE_AS_OF, significant, TOTAL_COST_CURRENT))
```

```{r slideshow={'slide_type': 'subslide'}}
subset(incidents, round(cost_1984) == 14956852)$narrative
```

```{r slideshow={'slide_type': 'subslide'}}
subset(incidents, round(cost_1984) == 10968391)$narrative
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
* The Chevron incident is the direct result of Hurricane Ivan in 2004:
    * https://en.wikipedia.org/wiki/Hurricane_Ivan
    * https://www.nytimes.com/2005/08/05/business/gulf-hurricanes-arelatest-kink-in-oil-chain.html
* Less information can be found on the damage to BP's offshore pipeline, but it is related to Hurrican Ivan also. The Hurricane cause earth movement, which for instance caused the Taylor Oil Spill. The Taylor platform used to belong to BP, so it seems likely that the same event also damaged a pipeline of BP in the area.
    * https://en.wikipedia.org/wiki/Taylor_oil_spill
<!-- #endregion -->

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## By commodity without offshore
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
pipelines_y_c <- pipelines %>%
    filter(on_offshore == "onshore") %>%
    filter(commodity %in% c("crude", "hvl", "rpp")) %>%
    group_by(year, commodity) %>%
    summarize(!!! oildata::pipelines_consolidation) %>%
    pivot_longer(cols = -c(year, commodity), names_to = "variable", values_to = "value")

glimpse(pipelines_y_c)
```

```{r}
overview_y_c <- pipelines_y_c %>%
    filter(variable %in% c("miles", "significant_incidents", "incidents_cost", "incidents_volume")) %>%
    ggplot(aes(x = year, y = value)) +
        facet_wrap(commodity ~ variable, scales = "free") +
        geom_line() +
        geom_smooth(method = "lm")

options(repr.plot.width=8, repr.plot.height=6)
```

```{r slideshow={'slide_type': 'subslide'}}
overview_y_c
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Relative to miles and volume transported
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
pipelines_y_c <- pipelines %>%
    filter(on_offshore == "onshore") %>%
    filter(commodity %in% c("crude", "hvl", "rpp")) %>%
    group_by(year, commodity) %>%
    summarize(!!! oildata::pipelines_consolidation) %>%
    mutate(cost_per_mile = incidents_cost/miles, 
           cost_per_volume = incidents_cost/estimate_volume_all, 
           volume_per_mile = incidents_volume/miles,
           volume_per_volume = incidents_volume/estimate_volume_all) %>%
    pivot_longer(cols = -c(year, commodity), names_to = "variable", values_to = "value")

glimpse(pipelines_y_c)
```

```{r slideshow={'slide_type': 'subslide'}}
standardized <- pipelines_y_c %>%
    filter(variable %in% c("cost_per_mile", "cost_per_volume", "volume_per_mile", "volume_per_volume")) %>%
    ggplot(aes(x = year, y = value)) +
        facet_wrap(commodity ~ variable, scales = "free") +
        geom_line() +
        geom_smooth(method = "lm")

options(repr.plot.width=8, repr.plot.height=6)
```

```{r slideshow={'slide_type': 'subslide'}}
standardized
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Detailed view
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
options(repr.plot.width=8, repr.plot.height=5)
```

```{r slideshow={'slide_type': 'subslide'}}
pipelines_y_c %>%
    filter(variable == "volume_per_volume") %>%
    ggplot(aes(x=year, y=value, color=commodity)) + 
        geom_line() +
        labs(y = "Gallons of oil spilled per gallon-mile transported")
```

```{r slideshow={'slide_type': 'subslide'}}
pipelines_y_c %>%
    filter(variable == "volume_per_volume") %>%
    ggplot(aes(x=year, y=value)) + 
        facet_wrap(~commodity) +
        geom_line() +
        geom_smooth(method = "lm") +
        labs(y = "Gallons of oil spilled per gallon-mile transported")
```

```{r slideshow={'slide_type': 'subslide'}}
pipelines_y_c %>%
    filter(variable == "cost_per_volume") %>%
    ggplot(aes(x=year, y=value)) + 
        facet_wrap(~commodity) +
        geom_line() +
        geom_smooth(method = "lm") +
        labs(y = "Cost of damage by oil spills in USD per gallon-mile transported")
```

```{r slideshow={'slide_type': 'subslide'}}
cost_graph <- pipelines_y_c %>%
    filter(variable == "cost_per_volume") %>%
    ggplot(aes(x=year, y=value)) + 
        facet_wrap(~commodity) +
        geom_line() +
        geom_smooth(method = "lm") +
        labs(y = "Cost of damage by oil spills in USD per gallon-mile transported") +
        scale_y_log10() +
        labs(title = "!!! Cost - log scaled !!!")
```

```{r slideshow={'slide_type': 'subslide'}}
cost_graph
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Prominent cases
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
largest_companies <- pipelines %>%
    filter(on_offshore == "onshore") %>%
    filter(commodity %in% c("rpp", "crude", "hvl")) %>%
    group_by(year, ID) %>%
    summarize(miles = sum(miles, na.rm = T)) %>%
    group_by(ID) %>%
    filter(miles == max(miles)) %>%
    ungroup() %>%
    top_n(6, miles) %>%
    arrange(desc(miles))

largest_companies %>%
    jupyter_styling(font_size = 14)
```

```{r}
add_company_names(31618)
```

```{r slideshow={'slide_type': 'skip'}}
operators <- pipelines %>%
    filter(ID %in% c("31618", "Energy Transfer Partners (Group)", "Kinder Morgan (Group)")) %>%
    mutate(ID = add_company_names(ID)) %>%
    filter(on_offshore == "onshore") %>%
    filter(commodity %in% c("crude", "hvl", "rpp")) %>%
    group_by(year, commodity, ID) %>%
    summarize(!!! oildata::pipelines_consolidation) %>%
    mutate(cost_per_mile = incidents_cost/miles, 
           cost_per_volume = incidents_cost/estimate_volume_all, 
           volume_per_mile = incidents_volume/miles,
           volume_per_volume = incidents_volume/estimate_volume_all) %>%
    pivot_longer(cols = -c(year, commodity, ID), names_to = "variable", values_to = "value")
glimpse(operators)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Volume spilled per volume transported
<!-- #endregion -->

```{r slideshow={'slide_type': 'skip'}}
crude <- operators %>%
    filter(variable %in% c("volume_per_volume")) %>%
    filter(commodity == "crude") %>%
    ggplot(aes(x=year, y=value)) +
        facet_wrap(~ ID) +
        geom_point() +
        geom_line() +
        scale_y_log10() +
        geom_smooth(method = "lm", se = F) +
        labs(title = "Crude oil", x = NULL, y = NULL)

hvl <- operators %>%
    filter(variable %in% c("volume_per_volume")) %>%
    filter(commodity == "hvl") %>%
    ggplot(aes(x=year, y=value)) +
        facet_wrap(~ ID) +
        geom_point() +
        geom_line() + 
        scale_y_log10() +
        geom_smooth(method = "lm", se = F) +
        labs(title = "Highly volatile liquid", x = NULL, y = NULL)

rpp <- operators %>%
    filter(variable %in% c("volume_per_volume")) %>%
    filter(commodity == "rpp") %>%
    ggplot(aes(x=year, y=value)) +
        facet_wrap(~ ID) +
        geom_point() +
        geom_line() + 
        scale_y_log10() +
        geom_smooth(method = "lm", se = F) +
        labs(title = "Refined petroleum products", x = NULL, y = NULL)
```

```{r slideshow={'slide_type': 'skip'}}
spill_volume_by_operator <- arrangeGrob(crude, hvl, rpp, ncol=1, left = "Gallons of oil spilled per gallon-mile transported")

options(repr.plot.width=8, repr.plot.height=6)
```

```{r slideshow={'slide_type': 'subslide'}}
grid::grid.draw(spill_volume_by_operator)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Total miles
<!-- #endregion -->

```{r slideshow={'slide_type': 'skip'}}
crude <- operators %>%
    filter(variable %in% c("miles")) %>%
    filter(commodity == "crude") %>%
    ggplot(aes(x=year, y=value)) +
        facet_wrap(~ ID) +
        geom_point() +
        geom_line() +
        scale_y_log10() +
        geom_smooth(method = "lm", se = F) +
        labs(title = "Crude oil", x = NULL, y = NULL)

hvl <- operators %>%
    filter(variable %in% c("miles")) %>%
    filter(commodity == "hvl") %>%
    ggplot(aes(x=year, y=value)) +
        facet_wrap(~ ID) +
        geom_point() +
        geom_line() + 
        scale_y_log10() +
        geom_smooth(method = "lm", se = F) +
        labs(title = "Highly volatile liquid", x = NULL, y = NULL)

rpp <- operators %>%
    filter(variable %in% c("miles")) %>%
    filter(commodity == "rpp") %>%
    ggplot(aes(x=year, y=value)) +
        facet_wrap(~ ID) +
        geom_point() +
        geom_line() + 
        scale_y_log10() +
        geom_smooth(method = "lm", se = F) +
        labs(title = "Refined petroleum products", x = NULL, y = NULL)
```

```{r slideshow={'slide_type': 'skip'}}
volume_transported_by_operator <- arrangeGrob(crude, hvl, rpp, ncol=1, left = "Miles of pipeline")

options(repr.plot.width=8, repr.plot.height=6)
```

```{r slideshow={'slide_type': 'subslide'}}
grid::grid.draw(volume_transported_by_operator)
```

```{r slideshow={'slide_type': 'subslide'}}
grid::grid.draw(spill_volume_by_operator)
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
### Using size argument
<!-- #endregion -->

```{r slideshow={'slide_type': 'skip'}}
operators2 <- operators %>%
    filter(variable %in% c("volume_per_volume", "miles")) %>%
    pivot_wider(id_cols = c(year, commodity, ID), names_from = "variable", values_from = "value")
```

```{r slideshow={'slide_type': 'skip'}}
crude <- operators2 %>%
    filter(commodity == "crude") %>%
    ggplot(aes(x=year, y=volume_per_volume, size=miles)) +
        facet_wrap(~ ID) +
        geom_point() +
        geom_line(aes(size=NULL)) +
        scale_y_log10() +
        geom_smooth(method = "lm", se = F) +
        labs(title = "Crude oil", x = NULL, y = NULL)

hvl <- operators2 %>%
    filter(commodity == "hvl") %>%
    ggplot(aes(x=year, y=volume_per_volume, size=miles)) +
        facet_wrap(~ ID) +
        geom_point() +
        geom_line(aes(size=NULL)) +
        scale_y_log10() +
        geom_smooth(method = "lm", se = F) +
        labs(title = "Highly volatile liquid", x = NULL, y = NULL)

rpp <- operators2 %>%
    filter(commodity == "rpp") %>%
    ggplot(aes(x=year, y=volume_per_volume, size=miles)) +
        facet_wrap(~ ID) +
        geom_point() +
        geom_line(aes(size=NULL)) +
        scale_y_log10() +
        geom_smooth(method = "lm", se = F) +
        labs(title = "Refined petroleum products", x = NULL, y = NULL)
```

```{r slideshow={'slide_type': 'skip'}}
volume_per_size <- arrangeGrob(crude, hvl, rpp, ncol=1, left = "Gallons of oil spilled per gallon-mile transported")

options(repr.plot.width=8, repr.plot.height=6)
```

```{r slideshow={'slide_type': 'subslide'}}
grid::grid.draw(volume_per_size)
```

<!-- #region {"slideshow": {"slide_type": "slide"}} -->
## Test using all volume vs specific volume
<!-- #endregion -->

```{r slideshow={'slide_type': 'subslide'}}
using_volume_specific <- pipelines %>%
    filter(on_offshore == "onshore") %>%
    filter(commodity %in% c("crude", "hvl", "rpp")) %>%
    group_by(year, commodity) %>%
    summarize(!!! oildata::pipelines_consolidation) %>%
    mutate(volume_per_volume_specific = (
               incidents_volume/estimate_volume_specific), 
           cost_per_volume_specific = (
               incidents_cost/estimate_volume_specific)) %>%
    pivot_longer(cols = -c(year, commodity), 
                 names_to = "variable", values_to = "value") %>%
    filter(variable %in% c("cost_per_volume_specific", 
                           "volume_per_volume_specific")) %>%
    ggplot(aes(x = year, y = value)) +
        facet_wrap(variable ~ commodity, scales = "free") +
        geom_line() +
        geom_smooth(method = "lm")

options(repr.plot.width=8, repr.plot.height=6)
```

```{r slideshow={'slide_type': 'subslide'}}
using_volume_specific
```

<!-- #region {"slideshow": {"slide_type": "subslide"}} -->
It looks very much the same as the previous graph, using volume_all.
<!-- #endregion -->
